CDSR Human Data/Modelling Paper

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK updated the text in the analysis section and resized the bottleneck1.jpg

\section{Human Studies - John}

@@

The study was designed to examine the following questions about spatial regions:
\begin{enumerate}
	\item Given a particular situation how well do people agree on the location and extent of a named context dependent spatial region?
	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named context dependent spatial region?
\end{enumerate}

To examine these questions, we developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).  Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment. Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli. The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and (3) the military engagement stimuli were always accompanied by the description \emph{safety}.

@@

The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region. Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied). Again, a variation across the between subject agreement on the different stimuli would provide evidence for the impact of object location and orientation on spatial semantics.

@@

%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
%\begin{equation*}
%\kappa = \frac{P(a)-P(ca)}{1-P(ca)}
%\end{equation*}
%where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators. 
In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region. In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.

@@

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain. For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli. In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli. For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sampl t-tests. There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points. Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli. In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. Table \ref{tab:flipped-point-likert-scores} lists the results of this trial. In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.

\begin{figure}
%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
\centerline{
\begin{tabular}{ccc}
\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
\end{tabular}
}
\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
\label{fig:locations}
\end{figure}

\begin{table}
\centerline{
\begin{tabular}{l c c c l}
\hline
Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
\\
\hline
Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
\\
\hline
Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
\\
\hline
\hline
\end{tabular}
}
\label{tab:point-likert-scores}
\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
\end{table}

>>>>

\section{Human Studies}

@@

The study was designed to examine the following questions about CDSRs:
\begin{enumerate}
	\item Given a particular situation how well do people agree on the location and extent of a named context dependent spatial region?
	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named context dependent spatial region?
\end{enumerate}

Both of these research questions address the fundamental claim that CDSRs are contextually defines. Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects. Research question 2 address the contextual nature of CDSRs. Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

To examine these questions, we developed two experimental tasks:
\begin{enumerate}
	\item Task 1: subjects were presented with a visual representation of a spatial scene and an linguistic description containing a CDSR and were tasked with drawing a polygon a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene or to select a no region opt
	\item Task 2: subjects were presented with a visual representation of a spatial scene where a particular location in the scene had been marked and an accompanying spatial description and were taked with rating on a five point likert scale how well the marked location fitted with their understanding of the CDSR.
\end{enumerate}

Both Task 1 and 2 addressed both research questions. Between subject agreement metrics can be computed for both Task 1 and Task 2 and, consequently, both tasks address research question 1. In particular, a variation in between subject agreement across the different contexts for a particular CDSR could be taken as evidence that some contexts are more amenable to a particular CDSR than others and hence CDSRs are contextually dependent. By shifting the focus of the analysis to examining the variation in responses for a given CDSR in context that differed only with respect to the location of objects both the results from both tasks can address research question 2. In the case of Task 1 the variation we will examine is the area of the polygons, which in Task 2 the variation we will examine is the scoring of the marked locations.

We developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).  Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment. These stimuli were used in both Task 1 and 2. Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli. The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and (3) the military engagement stimuli were always accompanied by the description \emph{safety}.

@@

The geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the width of the bottleneck region. We hoped this variation woudd result in a variation in the size of the polygons for these different visual contexts. Finally, the military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied). Again, a variation across the between subject agreement on the different stimuli would provide evidence for the impact of object location and orientation on spatial semantics.

@@

%Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes as is defined as follows:
%\begin{equation*}
%\kappa = \frac{P(a)-P(ca)}{1-P(ca)}
%\end{equation*}
%where $P(a)$ is the observed relative agreement between the annotators, and $P(ca)$ is the probability of a chance agreement between these to annotators.

%In the type 1 trials subjects drew a polygon on the visual stimuli that defined the location and extent of the relevant spatial region. In analysising the results of these trials we wished to investigate how well the subject's responses agreed with each other and, also, whether the location and orientation of objects in the scene effected the definition of spatial regions in the scene.

@@

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain. For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli. In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli. For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points. 
%There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points. Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli. In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. %Table \ref{tab:flipped-point-likert-scores} lists the results of this trial. 
In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli. In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.

%\begin{figure}
%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
%\centerline{
%\begin{tabular}{ccc}
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/classroom-circle-points.png}\caption{}\label{fig:classroom-points}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/bottleneck-2-points.png}\caption{}\label{fig:bottleneck-points}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/tanks-close-points.png}\caption{}\label{fig:tanks-points}\end{subfigure}\\
%\end{tabular}
%}
%\caption{The pre-selected locations for each domain used in the Type 2 Trials.}
%\label{fig:locations}
%\end{figure}

\begin{table}
\centerline{
\begin{tabular}{l c c c }
\hline
Domain & \multicolumn{3}{c}{Map/Bottleneck} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} \\
Stim ID & P1 & P2 & P3 \\
(g) & $1.1707$ & $\mathbf{4.6667}$ \\ % 1_1
(h) & $1.1707$ & $\mathbf{4.7561}$ \\ % 1_2
\\
\hline
Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} \\
Stim ID & P1 & P2 & P3 \\
(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ \\ % 2_1
(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ \\ % 2_6
(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ \\ %2_7
(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ \\ %2_2
(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ \\ % 2_3
(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ \\ % 2_4
(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ \\ % 2_5
\\
\hline
Domain & \multicolumn{3}{c}{Military Engagement/Safe Region} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} \\
Stim ID & P1 & P2 & P3\\
(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ \\ % 3_1
(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ \\ % 3_2
(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ \\ % 3_3
(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$\\ % 3_4
\\
\hline
\hline
\end{tabular}
}
\label{tab:point-likert-scores}
\caption{Average Likert Ratings per Point by Domain.}
\end{table}

%\begin{table}
%\centerline{
%\begin{tabular}{l c c c l}
%\hline
%Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
% ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
%(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
%(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
%\\
%\hline
%Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
% ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
%(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
%(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
%(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
%(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
%(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
%(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
%(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
%\\
%\hline
%Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
% ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
%Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
%(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
%(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
%(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
%(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
%\\
%\hline
%\hline
%\end{tabular}
%}
%\label{tab:point-likert-scores}
%\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
%\end{table}
